Variance reduction for Markov chains via martingale representations

### 2020-05-08 aap/Revision/vr-mcmc-revision-final.tex
update

Under assumptions {\bf (H1)} and {\bf (H2)} the Markov chain $(X^x_{p})_{p \in \nset_0}$ generated by ULA  is $V-$geometrically ergodic in the sense of \eqref{eq:v_ergodicity} with $V(x) = 1+|x|^2$ and $\rho\in (0,1) $ due to Lemma~\ref{lem:v_ergodicity}. It follows from Lemma~\ref{lem:var_bound}, that
\[
\PVar[\pi_n^N(f)]\lesssim \frac{1}{n(1-\rho^{1/2})} + \left(\frac{\rho^NV(x)+1}{n(1-\rho^{1/2})}\right)^2
\]
It follows from \eqref{eq:rho-ula} that $1-\rho^{1/2} \geq C\gamma$ with a constant $C$ not depending on $\gamma,N$ and $n$, thus yielding an estimate
\[
\PVar[\pi_n^N(f)] \lesssim \frac{1}{n\gamma} + \left(\frac{\rho^NV(x)+1}{n\gamma}\right)^2.
\]

@@

\begin{lem}
\label{lem:v_ergodicity}
 Assume  {\bf (H1)} and {\bf (H2)}. Then for $0 < \gamma < \overline{\gamma}= \min(\mathfrak{m} / 2L^2,1/L)$, and for any $x \in \rset^d$  it holds
$$
d_V(\delta_xQ_{\gamma}^{n},\pi_\gamma) \leq C\rho^{\gamma n} V(x)
$$
with $V(x) = 1 + |x|^2$ and some $0 < \rho < 1$ and $C > 0$.
%$$
%C = \left(1 + \exp{\left(-\frac{\mathfrak{m}\gamma}{2}\right)}\right)\left(1+\frac{\overline{b}}{(1-\varepsilon)(1-\exp{\left(-\frac{\mathfrak{m}\gamma}{2}\right)} - \frac{2b}{2+K^2})}\right),
%$$
%
%
%
%$$
%b = \gamma(2K_2^2(L+\mathfrak{m}) + d + \mathfrak{m}), \quad \overline{b} = b\exp{\left(-\frac{\mathfrak{m}\gamma}{2}\right)} + K^2 + 1, 
%$$
%
%\begin{eqnarray*}
%\varepsilon = 2\Phi\left(-K\sqrt{\left(1+\frac{1}{L}\right)\left(1 + 2L\right)}\right),
%\end{eqnarray*}
%
%\begin{eqnarray}
%\label{eq:rho-ula}
%\rho = \exp\left\{-\frac{\gamma m}{4}\frac{\log{(1-\varepsilon)}}{\log{(1-\varepsilon)} + \log{(\exp{\left(-\frac{\mathfrak{m}\gamma}{2}\right)} + \frac{2b}{K^2+2})}}\right\},
%\end{eqnarray}
%where $K_2 > 0$ is from Lemma~\ref{lem:quadratic_behaviour} and $K>0$ is a positive parameter such that 
%$$
%K^2 > 4\frac{2K_2^2(L+m) + d + m}{m}.
%$$
\end{lem}

>>>>

Under assumptions {\bf (H1)} and {\bf (H2)} the Markov chain $(X^x_{p})_{p \in \nset_0}$ generated by ULA  is $V-$geometrically ergodic in the sense of \eqref{eq:v_ergodicity} with $V(x) = 1+|x|^2$ due to Lemma~\ref{lem:v_ergodicity}. It follows from Lemma~\ref{lem:var_bound}, applied with $\rho^{\gamma}$ that
\[ 
\PVar[\pi_n^N(f)]\lesssim \frac{1}{n(1-\rho^{\gamma/2})} + \left(\frac{\rho^{\gamma N}V(x)+1}{n(1-\rho^{\gamma/2})}\right)^2
\]
Since $1-\rho^{\gamma/2} \geq C\gamma$ with a constant $C$ not depending on $\gamma,N$ and $n$, we end up with an estimate
\[
\PVar[\pi_n^N(f)] \lesssim \frac{1}{n\gamma} + \left(\frac{\rho^NV(x)+1}{n\gamma}\right)^2.
\]

@@

\begin{lem}
\label{lem:v_ergodicity}
 Assume  {\bf (H1)} and {\bf (H2)}. Then for $0 < \gamma < \overline{\gamma}= \min(\mathfrak{m} / 2L^2,1/L)$, and for any $x \in \rset^d$  it holds
\begin{equation}
\label{eq:rho-ula}
d_V(\delta_xQ_{\gamma}^{n},\pi_\gamma) \leq C\rho^{\gamma n} V(x)
\end{equation}
with $V(x) = 1 + |x|^2$ and some $0 < \rho < 1$ and $C > 0$.
%$$
%C = \left(1 + \exp{\left(-\frac{\mathfrak{m}\gamma}{2}\right)}\right)\left(1+\frac{\overline{b}}{(1-\varepsilon)(1-\exp{\left(-\frac{\mathfrak{m}\gamma}{2}\right)} - \frac{2b}{2+K^2})}\right),
%$$
%
%
%
%$$
%b = \gamma(2K_2^2(L+\mathfrak{m}) + d + \mathfrak{m}), \quad \overline{b} = b\exp{\left(-\frac{\mathfrak{m}\gamma}{2}\right)} + K^2 + 1, 
%$$
%
%\begin{eqnarray*}
%\varepsilon = 2\Phi\left(-K\sqrt{\left(1+\frac{1}{L}\right)\left(1 + 2L\right)}\right),
%\end{eqnarray*}
%
%\begin{eqnarray}
%\label{eq:rho-ula}
%\rho = \exp\left\{-\frac{\gamma m}{4}\frac{\log{(1-\varepsilon)}}{\log{(1-\varepsilon)} + \log{(\exp{\left(-\frac{\mathfrak{m}\gamma}{2}\right)} + \frac{2b}{K^2+2})}}\right\},
%\end{eqnarray}
%where $K_2 > 0$ is from Lemma~\ref{lem:quadratic_behaviour} and $K>0$ is a positive parameter such that 
%$$
%K^2 > 4\frac{2K_2^2(L+m) + d + m}{m}.
%$$
\end{lem}
